Non-word repetition in Yélî Dnye

### 2020-09-16 manuscript.tex
add comments for AC + fix fig IPA, table on length, and one ref

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[english,,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Non-word repetition in Yélî Dnye},
            pdfauthor={Alejandrina Cristia~\& Marisa Casillas},
            pdfkeywords={phonology, non-word repetition, development},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\setmainfont{Doulos SIL} % Set main font to Doulos SIL
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi

@@

\hypertarget{todo-middy}{%
\subsection{TODO Middy}\label{todo-middy}}

\begin{itemize}
\tightlist
\item
  look through this manuscript for bits of text to be written by \textbf{mc}
\item
  do you think Figure 2 should be done with phonetic symbols? If yes, could you try to implement that?
\item
  double check demo -- probably start with the xlsx version that has the children's first names, date of birth, and date of test. I looked again at my local files (still haven't checked the labs' ones) and I'm surprised I have maternal education for more children than I have birth order for\ldots{} If you correct this, to generate the csv I also round the age, assign age.rounded to reported age (in years) when DOB is missing, remove first names, DOB, DOTest and age.precise before pushing to github
\item
  double check my frequency entries in segments.xlsx using the search function in \url{https://phoible.org/parameters}; take a look at \url{http://phoible.github.io/conventions/} in case you see something about double articulation (I didn't find tp or lBj)
\item
  double check phonological implementation of stimuli in stimuli.xlsx
\item
  for both of the above, if you make any changes, they need to be re-saved as UTF16 txt file, and then converted (see remember above)
\item
  double check also stim\_seg\_freq -- why are some items still with NA freq?
\item
  draft discussion
\end{itemize}

\hypertarget{todo-alex}{%
\subsection{TODO Alex}\label{todo-alex}}

@@

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

@@

\hypertarget{yuxe9luxee-dnye-phonological-overview}{%
\subsubsection{Yélî Dnye phonological overview}\label{yuxe9luxee-dnye-phonological-overview}}

Yélî Dnye is an isolate language (presumed Papuan) spoken by approximately 7,000 people residing on Rossel Island, which is found at the far end of the Louisiade Archipelago in Milne Bay Province, Papua New Guinea. The Yélî sound system, much like its baroque grammatical system (Levinson, n.d.), is unlike any other in the region.

With only four primary places of articulation (bilabial, alveolar, post-alveolar, and velar) and no voicing contrasts, the phonological inventory is remarkably packed with acoustically similar segments. The core oral stop set includes both singleton (/p/, /t/, /ṭ/, and /k/) and doubly-articulated (/tp/, /ṭp/, /kp/) segments, with full nasal equivalents (/m/, /n/, /ṇ/, /ŋ/, /nm/, /ṇm/, /ŋm/), and with a substantial portion of these able to be contrastively pre-nasalized or nasally released (/mp/, /nt/, /ṇṭ/, /ŋk/, /nmtp/, /ṇmṭp/, /ŋmkp/, /ṭṇ/, /kŋ/, /ṭpṇm/, /kpŋm/). A large number of this combinatorial set can further be contrastively labialized or palatalized on release, or both (e.g., /pʲ/, /pʷ/, /pʲʷ/; /tpʲ/; /ṇmḍbʲ/; see Levinson (n.d.) for details). The consonantal inventory also includes a number of non-nasal continuants (/w/, /j/, /ɣ/, /l/, /βʲ/, /lʲ/, /lβʲ/). Vowels in Yélî Dnye may be oral or nasal, short or long. The 10 oral vowel qualities, which span four levels of vowel height, (/i/, /ɯ/, /u/, /e/, /o/, /ə/, /ɛ/, /ɔ/, /æ/, /ɑ/) can be produced as short or long vowels, with 7 of these able to appear as short and long nasal vowels as well (/i/, /u/, /ə/, /ɛ/, /ɔ/, /æ/, /ɑ/).

In total, Yélî Dnye uses 90 distinctive segments (93 when including consonants that are extremely rarely used), far outstripping the phonemic inventory size of other documented Papuan languages (Levinson (n.d.); Foley 1986) and includes at least two contrasts not yet been documented elsewhere {[}labial-coronal double-articulations with dental vs.~post-alveolar coronal placement in both oral and nasal stops, Maddieson \& Levinson 2001; Ladefoged and Maddieson 1996{]}.

Most words in Yélî Dnye are bisyllabic (\textasciitilde{}50\%), with monosyllabic words (\textasciitilde{}40\%) appearing most commonly after that, and with tri-and-above syllabic words appearing least frequently (\textasciitilde{}10\%; based on \textgreater{}5800 lexemes in the current Yélî Dnye dictionary, Levinson, n.d.). The vast majority of syllables use a CV format. A small portion of the lexicon features words with a final closed syllable, limited to codas of -/m/, -/p/, or -/j/ (e.g., `ndap' (Spondylus shell) /ṇṭæp/). However, in spontaneous speech an epenthetic /ɯ/ is often inserted after word-final coda, yielding a CVCV structure instead (e.g., `ndapî' /`ṇṭæ.pɯ/). This process is used frequently with English loan words that have a coda (e.g., 'ponî' (phone) /`pɔ.nɯ/). An even smaller portion of the lexicon features words starting with a vowel (e.g., 'ala' (here) /æ.'læ/), but these are limited to /æ/-. Finally, the lexicon features a handful of single-vowel grammatical morphemes (see Levinson (n.d.) for details).

While all consonants can appear in the first syllable of the word, the second syllable onset consonant in bisyllabic roots is almost exclusively drawn from a subset of the simplex consonants (in order of descending frequency: /m/, /ṭ/, /n/, /p/, /l/, /w/, /j/, /pʷ/, /ŋ/, /k/, /t/, /ɣ/, /mʷ/). In monosyllabic stems, oral vowels (short: 43\% and long: 31\%) are more common than nasal ones (short: 4\% and long: 22\%), with the first syllable of bisyllabic stems following a similar pattern. Most vowels in the second syllable of bisyllabic stems are also short and oral, with a potential preference for rounding harmony between the first and second vowels. Stress typically falls on the first syllable of bisyllabic words, on the first and third syllables of tetrasyllabic words, and on the first or second syllable of trisyllabic words, depending on its vocalic features (see Levinson (n.d.) for details).

\hypertarget{the-yuxe9luxee-community}{%
\subsubsection{The Yélî community}\label{the-yuxe9luxee-community}}

@@

Most speakers of the language grow up speaking it monolingually until they begin attending school around the age of 7;
school instruction is in English. While monolingual Yélî Dnye upbringing is common, multilingual families are not unusual, particularly in the region around the Catholic Mission---the same region in which the current data were collected---, where there is a higher incidence of married-in mothers from other islands (Brown \& Casillas, n.d.). Children in these multilingual families grow up speaking Yélî Dnye and English, Tok Pisin, and/or some other language(s) from the region.

@@

Despite restricted outside contact, formal education is a priority for many Yélî families. A recent study surveying more than 40 parents of young children in this region of the island found that nearly all parents completed 6 or more years of education, with 65\% of mothers and 49\% of fathers completing the full range of locally available education (\textasciitilde{}9 years), with around half of those parents having pursued an additional 2+ years of education on other islands in the region or on mainland PNG (Casillas, Brown, \& Levinson, 2020). While there are a handful of local schools around the island, these may be well out of walking distance for many children (i.e., more than 1 hour on foot or by canoe each day), and so it is very common for households to billet school children during the weekdays for months at a time (or for whole months if the child's home is very far). Combined with practices of collective child guardianship within their close-knit hamlets, the practical outcome for researchers is that adult consent can often come from a combination of aunts, uncles, adult cousins, and grandparents standing in for the child's parents---to this child assent is also culturally pertinent, as child independence is encouraged and respected from toddlerhood onward (Brown \& Casillas, n.d.).\footnote{Note that in a few cases this does mean that precise details such as child date of birth, parent age in years, and parent education was only able to be approximated as the parents were residing on another part of the island.} Finally, as mentioned, the language of schooling in this region is English, and so the majority of teens and adults on the island speak some English---many speak it very fluently.

While work on Yélî language development is growing (e.g., Brown, 2011, 2014; Brown \& Casillas, n.d.; Casillas et al., 2020; Liszkowski, Brown, Callaghan, Takada, \& de Vos, 2012), our knowledge of children's linguistic development is quite limited, and research into their phonological development in particular has only just begun {[}e.g., Bram's thesis; MC's CFH data{]}. What we do know about the early language environment is that, while child-directed speech is relatively infrequent throughout the day (Casillas et al., 2020), shared caregiving practices and a near-universal fondness for social interaction with young children ensure that speech directed to children comes from all types of speakers: women, men, and other children (Bunce et al., n.d.; Casillas et al., 2020). While speech from women predominates in young children's speech environments on Rossel, as it does elsewhere (Bergelson et al., 2019; Bunce et al., n.d.), there is a significant and marked increase in child-directed speech from other children as infants get older (Bunce et al., n.d.; Casillas et al., 2020). This increase in child-directed speech from other children is attributed to the fact that, starting around age 2, children often spend large swaths of the day playing, swimming, and foraging for fruit, nuts, and shellfish in large (\textasciitilde{}10 people) independent and mixed-age child play groups (Brown \& Casillas, n.d.; Casillas et al., 2020).

\hypertarget{brief-review-of-nwr-for-our-purposes}{%
\subsubsection{Brief review of NWR for our purposes}\label{brief-review-of-nwr-for-our-purposes}}

@@

A second list of candidate items was generated in 2019 by selecting complex consonants and systematically crossing them with all the vowels in the Yélî inventory to produce consonant-vowel monosyllables. As before, items were automatically excluded if they appeared in the dictionary. Additionally, since hearing vowel length in monosyllables in isolation is challenging, any item that had a short/long real word neighbor was filtered out. Since the phonology and phonetics of Yélî is still in the process of being described (Levinson, n.d.), there could have been undocumented constraints that rendered items illegal. Therefore, we made sure that the precise consonant-vowel sequence occurred in some real word in the dictionary (i.e., that there was a longer word included the monosyllable as a subsequence). These candidates were presented to one informant, for a final check that they did not mean anything. Together with the 2018 selection, they were recorded using a headset \textbf{XXX }mc** please fill in** and an Olympus \textbf{XXX }mc** please fill in** from the written form presented together with the same item orally (by the second author). The complete recorded list was finally presented to two more informants, who could repeat all the items and who confirmed there were no real words. Even so, there was one monosyllable that was often identified as a real word (intended ``yî'' /yXX/; identified as ``yi'' /yi/, \emph{tree}). This item is removed from analyses below.

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/Fig2-xling_freq-1.pdf}
\caption{(\#fig:Fig2-xling\_freq)NWR scores for individual target words as a function of the average frequency with which each phone is found across languages.}
\end{figure}

We next checked whether this may actually be due to frequency of the sounds within the language. We estimated frequency of the phones present in the stimuli in a corpus of child-centered recordings ({\textbf{???}}), and applied the natural logarithm. As before, unattested sounds were not considered (i.e., they were declared NA so that they do not count for analyses). Phone frequency estimated from this corpus were correlated with cross-linguistic phone frequencies (r(23)=0.65, =0). Therefore, we fit another mixed logistic regression, this time declaring as fixed effects scaled cross-linguistic and corpus frequencies (averaged across all attested phones within each stimulus item), in addition to age. As before, the model contained random slopes for both child ID and target. In this model, both cross-linguistic frequency (ß = 0.51, SE ß = 0.25, p = 0.04) and age (ß = 0.33, SE ß = 0.13, p = 0.01) were significant predictors of word-level NWR scores, but corpus frequency (ß = 0.39, SE ß = 0.24, p = 0.11) was not.

@@

\hypertarget{nwr-scores-as-a-function-of-item-length}{%
\subsubsection{NWR scores as a function of item length}\label{nwr-scores-as-a-function-of-item-length}}

@@

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:tablength}NWR measured in whole-word scores, phoneme-based scores, and normalized Levenshtein Distance, separately for the four stimuli lengths.}

\begin{tabular}{lll}
\toprule
Word & \multicolumn{1}{c}{Phoneme} & \multicolumn{1}{c}{NLD}\\
\midrule
47 (22) & 100 (0) & 0 (0)\\
79 (22) & 100 (0) & 0 (0)\\
78 (19) & 100 (0) & 0 (0)\\
75 (32) & 100 (0) & 0 (0)\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

\hypertarget{factor-structuring-individual-variation}{%
\subsubsection{Factor structuring individual variation}\label{factor-structuring-individual-variation}}

@@

\hypertarget{acknowledgments}{%
\subsection{Acknowledgments}\label{acknowledgments}}

@@

\hypertarget{references}{%
\section{References}\label{references}}

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-is08042009language}{}%
Action, C. (2009). Language impairment in a multilingual society: Linguistic patterns and the road to assessment. \emph{Brussels: COST Office. Available Online at: Http://Www. Bi-Sli. Org}.

\leavevmode\hypertarget{ref-balladares2016socio}{}%
Balladares, J., Marshall, C., \& Griffiths, Y. (2016). Socio-economic status affects sentence repetition, but not non-word repetition, in Chilean preschoolers. \emph{First Language}, \emph{36}(3), 338--351. \url{https://doi.org/10.1177/0142723715626067}

\leavevmode\hypertarget{ref-bergelsoncasillas2019what}{}%
Bergelson, E., Casillas, M., Soderstrom, M., Seidl, A., Warlaumont, A. S., \& Amatuni, A. (2019). What do North American babies hear? A large-scale cross-corpus analysis. \emph{Developmental Science}, \emph{22}(1), e12724. \url{https://doi.org/10.1111/desc.12724}

\leavevmode\hypertarget{ref-bowey2001nonword}{}%
Bowey, J. A. (2001). Nonword repetition and young children's receptive vocabulary: A longitudinal study. \emph{Applied Psycholinguistics}, \emph{22}(3), 441--469.

\leavevmode\hypertarget{ref-brandeker2015language}{}%
Brandeker, M., \& Thordardottir, E. (2015). Language exposure in bilingual toddlers: Performance on nonword repetition and lexical tasks. \emph{American Journal of Speech-Language Pathology}, \emph{24}(2), 126--138.

\leavevmode\hypertarget{ref-brown2011cultural}{}%
Brown, P. (2011). The cultural organization of attention. In A. Duranti, E. Ochs, \& and Bambi B Schieffelin (Eds.), \emph{Handbook of Language Socialization} (pp. 29--55). Malden, MA: Wiley-Blackwell.

\leavevmode\hypertarget{ref-brown2014interactional}{}%
Brown, P. (2014). The interactional context of language learning in Tzeltal. In I. Arnon, M. Casillas, C. Kurumada, \& B. Estigarribia (Eds.), \emph{Language in interaction: Studies in honor of Eve V. Clark} (pp. 51--82). Amsterdam, NL: John Benjamins.

\leavevmode\hypertarget{ref-brownIPchildrearing}{}%
Brown, P., \& Casillas, M. (n.d.). Childrearing through social interaction on Rossel Island, PNG. In A. J. Fentiman \& M. Goody (Eds.), \emph{Esther Goody revisited: Exploring the legacy of an original inter-disciplinarian} (pp. XX--XX). New York, NY: Berghahn.

>>>>

\documentclass[english,,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Non-word repetition in Yélî Dnye},
            pdfauthor={Alejandrina Cristia~\& Marisa Casillas},
            pdfkeywords={phonology, non-word repetition, development},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\setmainfont{Doulos SIL} % Set main font to Doulos SIL

@@

\subsection{TODO Middy}\label{todo-middy}

\begin{itemize}
\tightlist
\item
  look through this manuscript for bits of text to be written by
  \textbf{mc}
\item
  double check demo -- probably start with the xlsx version that has the
  children's first names, date of birth, and date of test. I looked
  again at my local files (still haven't checked the labs' ones) and I'm
  surprised I have maternal education for more children than I have
  birth order for\ldots{} If you correct this, to generate the csv I
  also round the age, assign age.rounded to reported age (in years) when
  DOB is missing, remove first names, DOB, DOTest and age.precise before
  pushing to github
\item
  double check my frequency entries in segments.xlsx using the search
  function in \url{https://phoible.org/parameters}; take a look at
  \url{http://phoible.github.io/conventions/} in case you see something
  about double articulation (I didn't find tp or lBj)
\item
  double check also stim\_seg\_freq -- why are some items still with NA
  freq?
\item
  draft discussion
\end{itemize}

\subsection{TODO Alex}\label{todo-alex}

@@

\subsection{Introduction}\label{introduction}

@@

\subsubsection{Yélî Dnye phonological
overview}\label{yuxe9luxee-dnye-phonological-overview}

Yélî Dnye is an isolate language (presumed Papuan) spoken by
approximately 7,000 people residing on Rossel Island, which is found at
the far end of the Louisiade Archipelago in Milne Bay Province, Papua
New Guinea. The Yélî sound system, much like its baroque grammatical
system (Levinson, in preparation), is unlike any other in the region.

With only four primary places of articulation (bilabial, alveolar,
post-alveolar, and velar) and no voicing contrasts, the phonological
inventory is remarkably packed with acoustically similar segments. The
core oral stop set includes both singleton (/p/, /t/, /ṭ/, and /k/) and
doubly-articulated (/tp/, /ṭp/, /kp/) segments, with full nasal
equivalents (/m/, /n/, /ṇ/, /ŋ/, /nm/, /ṇm/, /ŋm/), and with a
substantial portion of these able to be contrastively pre-nasalized or
nasally released (/mp/, /nt/, /ṇṭ/, /ŋk/, /nmtp/, /ṇmṭp/, /ŋmkp/, /ṭṇ/,
/kŋ/, /ṭpṇm/, /kpŋm/). A large number of this combinatorial set can
further be contrastively labialized or palatalized on release, or both
(e.g., /pʲ/, /pʷ/, /pʲʷ/; /tpʲ/; /ṇmḍbʲ/; see Levinson (in preparation)
for details). The consonantal inventory also includes a number of
non-nasal continuants (/w/, /j/, /ɣ/, /l/, /βʲ/, /lʲ/, /lβʲ/). Vowels in
Yélî Dnye may be oral or nasal, short or long. The 10 oral vowel
qualities, which span four levels of vowel height, (/i/, /ɯ/, /u/, /e/,
/o/, /ə/, /ɛ/, /ɔ/, /æ/, /ɑ/) can be produced as short or long vowels,
with 7 of these able to appear as short and long nasal vowels as well
(/i/, /u/, /ə/, /ɛ/, /ɔ/, /æ/, /ɑ/).

In total, Yélî Dnye uses 90 distinctive segments (93 when including
consonants that are extremely rarely used), far outstripping the
phonemic inventory size of other documented Papuan languages (Levinson
(in preparation); Foley 1986) and includes at least two contrasts not
yet been documented elsewhere {[}labial-coronal double-articulations
with dental vs.~post-alveolar coronal placement in both oral and nasal
stops, Maddieson \& Levinson 2001; Ladefoged and Maddieson 1996{]}.

Most words in Yélî Dnye are bisyllabic (\textasciitilde{}50\%), with
monosyllabic words (\textasciitilde{}40\%) appearing most commonly after
that, and with tri-and-above syllabic words appearing least frequently
(\textasciitilde{}10\%; based on \textgreater{}5800 lexemes in the
current Yélî Dnye dictionary, Levinson, in preparation). The vast
majority of syllables use a CV format. A small portion of the lexicon
features words with a final closed syllable, limited to codas of -/m/,
-/p/, or -/j/ (e.g., `ndap' (Spondylus shell) /ṇṭæp/). However, in
spontaneous speech an epenthetic /ɯ/ is often inserted after word-final
coda, yielding a CVCV structure instead (e.g., `ndapî' /`ṇṭæ.pɯ/). This
process is used frequently with English loan words that have a coda
(e.g., 'ponî' (phone) /`pɔ.nɯ/). An even smaller portion of the lexicon
features words starting with a vowel (e.g., 'ala' (here) /æ.'læ/), but
these are limited to /æ/-. Finally, the lexicon features a handful of
single-vowel grammatical morphemes (see Levinson (in preparation) for
details).

While all consonants can appear in the first syllable of the word, the
second syllable onset consonant in bisyllabic roots is almost
exclusively drawn from a subset of the simplex consonants (in order of
descending frequency: /m/, /ṭ/, /n/, /p/, /l/, /w/, /j/, /pʷ/, /ŋ/, /k/,
/t/, /ɣ/, /mʷ/). In monosyllabic stems, oral vowels (short: 43\% and
long: 31\%) are more common than nasal ones (short: 4\% and long: 22\%),
with the first syllable of bisyllabic stems following a similar pattern.
Most vowels in the second syllable of bisyllabic stems are also short
and oral, with a potential preference for rounding harmony between the
first and second vowels. Stress typically falls on the first syllable of
bisyllabic words, on the first and third syllables of tetrasyllabic
words, and on the first or second syllable of trisyllabic words,
depending on its vocalic features (see Levinson (in preparation) for
details).

\subsubsection{The Yélî community}\label{the-yuxe9luxee-community}

@@

Most speakers of the language grow up speaking it monolingually until
they begin attending school around the age of 7; school instruction is
in English. While monolingual Yélî Dnye upbringing is common,
multilingual families are not unusual, particularly in the region around
the Catholic Mission---the same region in which the current data were
collected---, where there is a higher incidence of married-in mothers
from other islands (Brown \& Casillas, in press). Children in these
multilingual families grow up speaking Yélî Dnye and English, Tok Pisin,
and/or some other language(s) from the region.

@@

Despite restricted outside contact, formal education is a priority for
many Yélî families. A recent study surveying more than 40 parents of
young children in this region of the island found that nearly all
parents completed 6 or more years of education, with 65\% of mothers and
49\% of fathers completing the full range of locally available education
(\textasciitilde{}9 years), with around half of those parents having
pursued an additional 2+ years of education on other islands in the
region or on mainland PNG (Casillas, Brown, \& Levinson, 2020). While
there are a handful of local schools around the island, these may be
well out of walking distance for many children (i.e., more than 1 hour
on foot or by canoe each day), and so it is very common for households
to billet school children during the weekdays for months at a time (or
for whole months if the child's home is very far). Combined with
practices of collective child guardianship within their close-knit
hamlets, the practical outcome for researchers is that adult consent can
often come from a combination of aunts, uncles, adult cousins, and
grandparents standing in for the child's parents---to this child assent
is also culturally pertinent, as child independence is encouraged and
respected from toddlerhood onward (Brown \& Casillas, in
press).\footnote{Note that in a few cases this does mean that precise
  details such as child date of birth, parent age in years, and parent
  education was only able to be approximated as the parents were
  residing on another part of the island.} Finally, as mentioned, the
language of schooling in this region is English, and so the majority of
teens and adults on the island speak some English---many speak it very
fluently.

While work on Yélî language development is growing (e.g., Brown, 2011,
2014; Brown \& Casillas, in press; Casillas et al., 2020; Liszkowski,
Brown, Callaghan, Takada, \& de Vos, 2012), our knowledge of children's
linguistic development is quite limited, and research into their
phonological development in particular has only just begun {[}e.g.,
Bram's thesis; MC's CFH data{]}. What we do know about the early
language environment is that, while child-directed speech is relatively
infrequent throughout the day (Casillas et al., 2020), shared caregiving
practices and a near-universal fondness for social interaction with
young children ensure that speech directed to children comes from all
types of speakers: women, men, and other children (Bunce et al., under
review; Casillas et al., 2020). While speech from women predominates in
young children's speech environments on Rossel, as it does elsewhere
(Bergelson et al., 2019; Bunce et al., under review), there is a
significant and marked increase in child-directed speech from other
children as infants get older (Bunce et al., under review; Casillas et
al., 2020). This increase in child-directed speech from other children
is attributed to the fact that, starting around age 2, children often
spend large swaths of the day playing, swimming, and foraging for fruit,
nuts, and shellfish in large (\textasciitilde{}10 people) independent
and mixed-age child play groups (Brown \& Casillas, in press; Casillas
et al., 2020).

\subsubsection{Brief review of NWR for our
purposes}\label{brief-review-of-nwr-for-our-purposes}

@@

A second list of candidate items was generated in 2019 by selecting
complex consonants and systematically crossing them with all the vowels
in the Yélî inventory to produce consonant-vowel monosyllables. As
before, items were automatically excluded if they appeared in the
dictionary. Additionally, since hearing vowel length in monosyllables in
isolation is challenging, any item that had a short/long real word
neighbor was filtered out. Since the phonology and phonetics of Yélî is
still in the process of being described (Levinson, in preparation),
there could have been undocumented constraints that rendered items
illegal. Therefore, we made sure that the precise consonant-vowel
sequence occurred in some real word in the dictionary (i.e., that there
was a longer word included the monosyllable as a subsequence). These
candidates were presented to one informant, for a final check that they
did not mean anything. Together with the 2018 selection, they were
recorded using a headset \textbf{XXX }mc** please fill in** and an
Olympus \textbf{XXX }mc** please fill in** from the written form
presented together with the same item orally (by the second author). The
complete recorded list was finally presented to two more informants, who
could repeat all the items and who confirmed there were no real words.
Even so, there was one monosyllable that was often identified as a real
word (intended ``yî'' /yXX/; identified as ``yi'' /yi/, \emph{tree}).
This item is removed from analyses below.

@@

\begin{figure}[!t]

{\centering \includegraphics[width=0.65\linewidth]{nwr.by.freq.ITEM}

\caption{NWR scores for individual target words as a function of the average frequency with which each phone is found across languages.}\label{fig:Fig2-xling-freq}
\end{figure}

We next checked whether this may actually be due to frequency of the
sounds within the language. We estimated frequency of the phones present
in the stimuli in a corpus of child-centered recordings (Casillas et
al., 2020), and applied the natural logarithm. As before, unattested
sounds were not considered (i.e., they were declared NA so that they do
not count for analyses). Phone frequency estimated from this corpus were
correlated with cross-linguistic phone frequencies (r(23)=0.65, =0).
Therefore, we fit another mixed logistic regression, this time declaring
as fixed effects scaled cross-linguistic and corpus frequencies
(averaged across all attested phones within each stimulus item), in
addition to age. As before, the model contained random slopes for both
child ID and target. In this model, both cross-linguistic frequency (ß =
0.51, SE ß = 0.25, p = 0.04) and age (ß = 0.33, SE ß = 0.13, p = 0.01)
were significant predictors of word-level NWR scores, but corpus
frequency (ß = 0.39, SE ß = 0.24, p = 0.11) was not.

@@

\subsubsection{NWR scores as a function of item
length}\label{nwr-scores-as-a-function-of-item-length}

@@

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:tablength}NWR measured in whole-word scores, phoneme-based scores, and normalized Levenshtein Distance, separately for the four stimuli lengths.}
\begin{tabular}{lll}
\toprule
Word & \multicolumn{1}{c}{Phoneme} & \multicolumn{1}{c}{NLD}\\
\midrule
47 (22) & 100 (0) & 0 (0)\\
79 (22) & 100 (0) & 0 (0)\\
78 (19) & 100 (0) & 0 (0)\\
75 (32) & 100 (0) & 0 (0)\\
\bottomrule
\end{tabular}
\end{threeparttable}
\end{center}
\end{table}

\subsubsection{Factor structuring individual
variation}\label{factor-structuring-individual-variation}

@@

\subsection{Acknowledgments}\label{acknowledgments}

@@

\section{References}\label{references}

@@

\hypertarget{refs}{}
\hypertarget{ref-is08042009language}{}
Action, C. (2009). Language impairment in a multilingual society:
Linguistic patterns and the road to assessment. \emph{Brussels: COST
Office. Available Online at: Http://Www. Bi-Sli. Org}.

\hypertarget{ref-balladares2016socio}{}
Balladares, J., Marshall, C., \& Griffiths, Y. (2016). Socio-economic
status affects sentence repetition, but not non-word repetition, in
Chilean preschoolers. \emph{First Language}, \emph{36}(3), 338--351.
doi:\href{https://doi.org/10.1177/0142723715626067}{10.1177/0142723715626067}

\hypertarget{ref-bergelsoncasillas2019what}{}
Bergelson, E., Casillas, M., Soderstrom, M., Seidl, A., Warlaumont, A.
S., \& Amatuni, A. (2019). What do North American babies hear? A
large-scale cross-corpus analysis. \emph{Developmental Science},
\emph{22}(1), e12724.
doi:\href{https://doi.org/10.1111/desc.12724}{10.1111/desc.12724}

\hypertarget{ref-bowey2001nonword}{}
Bowey, J. A. (2001). Nonword repetition and young children's receptive
vocabulary: A longitudinal study. \emph{Applied Psycholinguistics},
\emph{22}(3), 441--469.

\hypertarget{ref-brandeker2015language}{}
Brandeker, M., \& Thordardottir, E. (2015). Language exposure in
bilingual toddlers: Performance on nonword repetition and lexical tasks.
\emph{American Journal of Speech-Language Pathology}, \emph{24}(2),
126--138.

\hypertarget{ref-brown2011cultural}{}
Brown, P. (2011). The cultural organization of attention. In A. Duranti,
E. Ochs, \& and Bambi B Schieffelin (Eds.), \emph{Handbook of Language
Socialization} (pp. 29--55). Malden, MA: Wiley-Blackwell.

\hypertarget{ref-brown2014interactional}{}
Brown, P. (2014). The interactional context of language learning in
Tzeltal. In I. Arnon, M. Casillas, C. Kurumada, \& B. Estigarribia
(Eds.), \emph{Language in interaction: Studies in honor of Eve V. Clark}
(pp. 51--82). Amsterdam, NL: John Benjamins.

\hypertarget{ref-brownIPchildrearing}{}
Brown, P., \& Casillas, M. (in press). Childrearing through social
interaction on Rossel Island, PNG. In A. J. Fentiman \& M. Goody (Eds.),
\emph{Esther Goody revisited: Exploring the legacy of an original
inter-disciplinarian} (pp. XX--XX). New York, NY: Berghahn.

\hypertarget{ref-bunceURcrosscultural}{}
Bunce, J., Soderstrom, M., Bergelson, E., Rosemberg, C., Stein, A.,
Alam, F., \ldots{} Casillas, M. (under review). A cross-cultural
examination of young children's everyday language experiences.

\hypertarget{ref-casillas2020early}{}
Casillas, M., Brown, P., \& Levinson, S. C. (2020). Early language
experience in a papuan community. \emph{Journal of Child Language},
\emph{XX}, XX--XX.
